HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-17 paper/main.tex
fixed diagram

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{amsmath}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{url}
\usepackage{pgfplotstable}
\usepackage{hyperref}
\usepackage{color}
\usepackage{lipsum,adjustbox}

@@

\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
      inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
  \node (ROOT) [label={\#7}] [fill=black, circle] {}
  child {node [label={\#1}] {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
  child {node (H1) [label={\#8}] [fill=black,circle] {}
    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2) [label={\#9}] [fill=black,circle] {}
      child {node  {} edge from parent[draw=none]}
      child {node  {} edge from parent[draw=none]}
      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
      child {node [label={\#10}] [fill=black,circle] {}
        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
  \end{tikzpicture}
\caption{\label{fig:ucca_example}
  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
indices.}
\end{figure}

\begin{figure*}

  \begin{tikzpicture}
    [level 1/.style={sibling distance=7.5em},
      level 2/.style={sibling distance=5em},
      level 3/.style={sibling distance=5em},
      level 4/.style={sibling distance=3em}]
    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
  \node (ROOT) [label={B}] [fill=blue, circle] {}
  child {node (H3) [label={A}] [fill=yellow,circle] {}
    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
    child {node (H1) [label={[orange]:O}] [fill=orange,circle] {}
      child {node [label={}] {became} edge from parent node[left]  {}}
      child {node [label={}] {a} edge from parent node[left]  {}}
      child {node [label={}] {passion} edge from parent node[left]  {}}
    }
    child {node (H2) [label={A}] [fill=yellow,circle] {}
      child {node {} edge from parent[draw=none] {}}
      child {node {} edge from parent[draw=none] {}}
      child {node {} edge from parent[draw=none] {}}
      child {node {} edge from parent[draw=none] {}}
      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
    }
  }
  child {node {} edge from parent[draw=none] {}}
  child {node (H4) [label={A}] [fill=yellow,circle] {}
    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
  }
  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
  child {node (H5) [label={A}] [fill=yellow,circle] {}  
    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
  };
  \draw[dashed] (H5) -> (she) ;

  \com{
  child {node (H1) [label={\#8}] [fill=black,circle] {}
    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2) [label={\#9}] [fill=black,circle] {}
      child {node  {} edge from parent[draw=none]}
      child {node  {} edge from parent[draw=none]}
      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
      child {node [label={\#10}] [fill=black,circle] {}
        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
    }
  \end{tikzpicture}
\caption{\label{fig:ucca_example2}
  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
indices.}
\end{figure*}

@@

This section summarises the manual annotation procedure performed in order 
to compute the HUME measure. 
We mark the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$,
defined by its UCCA annotation, relative to $t$.
We note that UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation. This would allow monolingual annotators to take part in the
evaluation. On the downside, the translation would have to be aligned to the
reference and more importantly, we would risk inaccurate judgements due to
the naturally occurring differences between the source and its reference
translations.}.

@@

\begin{figure*}[t]
    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
	English paraphrase).}
    \label{fig:ucca-tree-hume}
\end{figure*}

@@

Human annotation of the semantic units requires first deciding whether
a unit is {\it atomic}, i.e., that it has no meaning-bearing
sub-units, or {\it structural}. In most cases, atomic units 
correspond to individual words, but they may also correspond to unanalyzable
multi-word expressions, i.e., expressions that do not have meaningful sub-units.
When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may either be
atomic or structural, and represent the relation between the sub-units.
For instance, a Scene unit represents the relation between the
Scene elements, such as the predicate (main relation), participants and adverbials.
Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.
We make a three-way distinction between translation qualities with structural
nodes, as opposed to the two-way distinction with atomic units, as lexical units
often can have slight errors, in which cases we do want to provide partial credit.

@@

For example, consider the sentence ``After graduation, Tom moved
to NYC'' as the source $s$, see \figref{fig:ucca_example}.
The atomic units are the individual words,
nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation $t$
is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
  be in English for presentation purposes. In our experiments, $t$ is in German,
  Czech, Polish or Romanian.}.
Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
expresses the relation between the graduation and moving Scenes, is marked Bad,
as their relation in $t$ is reversed.

Another example is given in \figref{fig:ucca-tree-hume}.
%``Golf became an obsession for his oldest daughter, she took daily lessons and became very good'',
The figure shows an annotation of a UCCA tree, and an example annotation for
each of the units.
We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
individually.\oa{I'm not sure I agree with the UCCA annotation;
also, translation annotation for ``she'' is missing. and for ``very good'' is wrong}

@@

\figref{fig:interface} presents the user interface of HUME. The sentence at the top
shows the complete MT system output. The source sentence appears at the top, and
the source translation directly below it. Underneath the source sentence appears
its list of semantic units. For each semantic unit, including the one for the
entire source sentence, there is a textbox, displaying the yield of that unit.
Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
The user is asked to select a label for each unit,
by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the left of the unit's box.

The interface presents, for each unit, the segment of the translation which aligns to it.
This allows the user, especially in long sentences, to focus her attention on the parts
mostly likely to be relevant for her judgement. As the alignments are automatically derived,
and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
it and make a judgement according to the full translation sentence, in cases where the alignment
is misleading.

Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
In the event that $a(u)$ is discontiguous in the text, we mark words between the left
and right boundaries of $a(u)$ which are not contained in it (intervening words)
in a smaller red font. Intervening words are likely change the meaning of the translation
of $u$, and thus should be paid attention to when considering whether the translation
is correct or not.

@@

\begin{figure*}[t]
    \includegraphics[width=1\textwidth]{ucca-tree}
    \caption{UCCA Tree with scenes}
    \label{ucca-tree}
\end{figure*}

>>>>

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{amsmath}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{url}
\usepackage{pgfplotstable}
%\usepackage{hyperref}
\usepackage{color}
\usepackage{lipsum,adjustbox}

@@

\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
    inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
  \node (ROOT) [fill=black, circle] {}
  child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
  child {node (H1)  [fill=black,circle] {}
    child {node {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2)  [fill=black,circle] {}
      child {node  {} edge from parent[draw=none]}
      child {node  {} edge from parent[draw=none]}
      child {node  (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node  {moved} edge from parent node[right] {$P$}}
      child {node  [fill=black,circle] {}
        child {node  {to} edge from parent node[right] {$R$}}
        child {node  {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
  \end{tikzpicture}
\caption{\label{fig:ucca_example}
  Sample UCCA annotation. Edge labels mark UCCA categories. }
\end{figure}

\begin{figure*}
  \begin{adjustbox}{trim=2.3cm 0 0 0}
  \begin{tikzpicture}
    [level 1/.style={sibling distance=7.5em},
      level 2/.style={sibling distance=5em},
      level 3/.style={sibling distance=5em},
      level 4/.style={sibling distance=3em}]
    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
  \node (ROOT) [label={B}] [fill=blue, circle] {}
  child {node (H3) [label={A}] [fill=yellow,circle] {}
    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
      child {node [label={}] {became} edge from parent node[left]  {}}
      child {node [label={}] {a} edge from parent node[left]  {}}
      child {node [label={}] {passion} edge from parent node[left]  {}}
    }
    child {node (H2) [label={A}] [fill=yellow,circle] {}
      child {node {} edge from parent[draw=none] {}}
      child {node {} edge from parent[draw=none] {}}
      child {node {} edge from parent[draw=none] {}}
      child {node {} edge from parent[draw=none] {}}
      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
    }
  }
  child {node {} edge from parent[draw=none] {}}
  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
  }
  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
  child {node (H5) [label={A}] [fill=yellow,circle] {}  
    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
  };
  \draw[dashed] (H5) -> (she) ;

  \com{
  child {node (H1) [label={\#8}] [fill=black,circle] {}
    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2) [label={\#9}] [fill=black,circle] {}
      child {node  {} edge from parent[draw=none]}
      child {node  {} edge from parent[draw=none]}
      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
      child {node [label={\#10}] [fill=black,circle] {}
        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
    }
  \end{tikzpicture}
  \end{adjustbox}
\caption{\label{fig:ucca_example2}
  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
indices.}
\end{figure*}

@@

This section summarises the manual annotation procedure performed in order 
to compute the HUME measure. 
We mark the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
We note that UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation. This would allow monolingual annotators to take part in the
evaluation. On the downside, the translation would have to be aligned to the
reference and more importantly, we would risk inaccurate judgements due to
the naturally occurring differences between the source and its reference
translations.}.

@@

\begin{figure*}[t]
    \includegraphics[width=1\textwidth]{ucca-tree-mteval.jpg}
    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
	English paraphrase).}
    \label{fig:ucca-tree-hume}
\end{figure*}

@@

Human annotation of the semantic units requires first deciding whether
a unit is {\it structural}, i.e., has meaning-bearing sub-units,
or {\it atomic}. In most cases, atomic units
correspond to individual words, but they may also correspond to unanalyzable
multi-word expressions, i.e., expressions that do not have meaningful sub-units
(e.g., idioms like ``takes two to tango'').
When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may either be
atomic or structural. The labels assigned to such nodes reflect the extent to which
the relation between the sub-units.
For instance, a Scene unit represents the relation between the
Scene elements, such as the predicate (main relation), participants and adverbials.
Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.
We make a three-way distinction between translation qualities with structural
nodes, as opposed to the two-way distinction with atomic units, as lexical units
often can have slight errors, in which cases we do want to provide partial credit.

@@

%For example, consider the sentence ``After graduation, Tom moved
%to NYC'' as the source $s$, see \figref{fig:ucca_example}.
%The atomic units are the individual words,
%nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation $t$
%is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
%  be in English for presentation purposes. In our experiments, $t$ is in German,
%  Czech, Polish or Romanian.}.
%Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
%Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
%is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
%as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
%expresses the relation between the graduation and moving Scenes, is marked Bad,
%as their relation in $t$ is reversed.

Another example is given in \figref{fig:ucca-tree-hume}.
%``Golf became an obsession for his oldest daughter, she took daily lessons and became very good'',
The figure shows an annotation of a UCCA tree, and an example annotation for
each of the units.
We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
individually.

@@

\figref{fig:interface} presents the user interface of HUME. 
The translation appears at the top, and
the source sentence directly below it. Underneath the source sentence appears
its list of semantic units. For each semantic unit, including the one for the
entire source sentence, there is a textbox, displaying the yield of that unit.
Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
The user is asked to select a label for each unit,
by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the left of the unit's box.

The interface presents, for each unit, the translation segment aligned with it.
This allows the user, especially in long sentences, to focus her attention on the parts
most likely to be relevant for her judgement. As the alignments are automatically derived,
and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
it and make a judgement according to the full translation sentence, in cases where the alignment
is misleading.

Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
In the event that $a(u)$ is discontiguous in the text, words between the left
and right boundaries of $a(u)$ which are not contained in it (intervening words)
are presented in a smaller red font.
Intervening words are likely change the meaning of the translation
of $u$, and thus should be paid attention to when considering whether the translation
is correct or not.

@@

\begin{figure*}[t]
    \includegraphics[width=1\textwidth]{ucca-tree.jpg}
    \caption{UCCA Tree with scenes}
    \label{ucca-tree}
\end{figure*}
